On the uniqueness and stability of dictionaries for sparse representation of noisy signals

### 2018-12-25 manuscripts/IEEE/final_submission.tex
final sub

\begin{problem}\label{InverseProblem}
Find a matrix $\mathbf{B}$ and $k$-sparse $\mathbf{\overline x}_1, \ldots, \mathbf{\overline x}_N \in \mathbb{R}^{\overline m}$ satisfying $\|\mathbf{z}_i - \mathbf{B}\mathbf{\overline x}_i\|_2 \leq \eta$ for all $i$.
\end{problem}

Our claim is that there exist very general conditions under which any matrix $\mathbf{B}$ that solves this problem is always ``close" to $\mathbf{A}$, and likewise for the sparse codes $\mathbf{\overline x}_i$ and $\mathbf{x}_i$. In fact, provided $n \geq \min(2k,m)$, in almost all cases (\ref{InverseProblem}) is well-posed (as per Hadamard \cite{Hadamard1902}) given enough data (Cor.~\ref{ProbabilisticCor}). Importantly, these explicit, universal guarantees apply without the imposition of \emph{any} constraints on the recovered matrix $\mathbf{B}$ (e.g. that it, too, satisfy any injectivity criteria), beyond a cap on the number $m$ of dictionary elements it should contain.

@@

To see how Def. \ref{maindef} is motivated by Prob. \ref{InverseProblem}, suppose that $Y$ has a stable $k$-sparse representation in $\mathbb{R}^m$ and fix $\delta_1, \delta_2$ to be the desired recovery accuracy in \eqref{def1}. Now, consider any dataset $Z$ generated as in \eqref{LinearModel} with $\eta \leq \frac{1}{2} \varepsilon(\delta_1, \delta_2)$. From the triangle inequality, it follows that any matrix $\mathbf{B}$ and $k$-sparse $\mathbf{\overline x}_1, \ldots, \mathbf{\overline x}_N \in \mathbb{R}^m$ solving Prob.~\ref{InverseProblem} are necessarily within $\delta_1, \delta_2$ of the original dictionary $\mathbf{A}$ and codes $\mathbf{x}_i$, respectively.

@@

\begin{problem}\label{OptimizationProblem}
Find $\mathbf{B}$ and \mbox{$\mathbf{\overline x}_1, \ldots, \mathbf{\overline x}_N \in \mathbb{R}^{\overline m}$} that solve:
\begin{align}\label{minsum}
\min \sum_{i = 1}^N \|\mathbf{\overline x}_i\|_0 \ \ 
\text{subject to} \ \ \|\mathbf{z}_i - \mathbf{B}\mathbf{\overline x}_i\|_2 \leq \eta, \ \text{for all $i$}.
\end{align}
\end{problem}
We then sketch proofs of extensions of Thm.~\ref{DeterministicUniquenessTheorem} to data and dictionaries drawn from a probability distribution (Thm.~\ref{robustPolythm} and Cor.~\ref{ProbabilisticCor}). Finally, in Sec.~\ref{Discussion}, we discuss both theoretical and practical applications of our main mathematical findings.

>>>>

\begin{problem}\label{InverseProblem}
Find a real $n \times \overline m$ matrix $\mathbf{B}$ and $k$-sparse $\mathbf{\overline x}_1, \ldots, \mathbf{\overline x}_N$ satisfying $\|\mathbf{z}_i - \mathbf{B}\mathbf{\overline x}_i\|_2 \leq \eta$ for all $i$.
\end{problem}

Our claim is that there exist very general conditions under which any such matrix $\mathbf{B}$ always contains a submatrix ``close" to some submatrix of $\mathbf{A}$, and similarly for the sparse codes $\mathbf{\overline x}_i$ and $\mathbf{x}_i$. In fact, provided $n \geq \min(2k,m)$, in almost all cases \ref{InverseProblem} is well-posed (as per Hadamard \cite{Hadamard1902}) given enough data (Cor.~\ref{ProbabilisticCor}). Importantly, these explicit, universal guarantees apply without the imposition of \emph{any} constraints on the recovered matrix $\mathbf{B}$ (e.g. that it, too, satisfy any injectivity criteria) beyond an upper bound on its number of columns, $\overline m$.

@@

To see how Def. \ref{maindef} is motivated by Prob. \ref{InverseProblem}, suppose that $Y$ has a stable $k$-sparse representation in $\mathbb{R}^m$ and fix $\delta_1, \delta_2$ to be the desired recovery accuracy in \eqref{def1}. Now, consider any dataset $Z$ generated as in \eqref{LinearModel} with $\eta \leq \frac{1}{2} \varepsilon(\delta_1, \delta_2)$. From the triangle inequality, it follows that any $n \times m$ matrix $\mathbf{B}$ and $k$-sparse $\mathbf{\overline x}_1, \ldots, \mathbf{\overline x}_N$ solving Prob.~\ref{InverseProblem} are necessarily within $\delta_1, \delta_2$ of the original dictionary $\mathbf{A}$ and codes $\mathbf{x}_i$, respectively.

@@

\begin{problem}\label{OptimizationProblem}
Find real $\mathbf{B}$ and \mbox{$\mathbf{\overline x}_1, \ldots, \mathbf{\overline x}_N$} that solve:
\begin{align}\label{minsum}
\min \sum_{i = 1}^N \|\mathbf{\overline x}_i\|_0 \ \
\text{subject to} \ \ \|\mathbf{z}_i - \mathbf{B}\mathbf{\overline x}_i\|_2 \leq \eta, \ \text{for all $i$}.
\end{align}
\end{problem}
We then sketch proofs of extensions of Thm.~\ref{DeterministicUniquenessTheorem} to data and dictionaries drawn from a probability distribution (Thm.~\ref{robustPolythm} and Cor.~\ref{ProbabilisticCor}). Finally, in Sec.~\ref{Discussion}, we discuss both theoretical and practical applications of our main mathematical findings.
